---
title: "the_escape_artist"
output: html_document
date: "2024-06-03"
---

Libraries
```{r}
library(ggthemes)
library(RColorBrewer)
library(extrafont)
library(grid)
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggplot2)
library(ggpattern)
library(gridExtra)
library(patchwork)
library(dplyr)
```

```{r}
escape <- read.csv("hoppers.csv")

str(escape)

escape$Location <- factor(escape$Location, 
                             levels = c("matched", "unmatched"), 
                             labels = c("Matched", "Unmatched"))

escape$Location <- factor(escape$Location, levels = c("Unmatched", "Matched"))
```

```{r}
is.factor(escape$Location)
escape$Location <- as.factor(escape$Location)
is.factor(escape$Location)

model <- lmer(Initiation_distance ~ Location + Flight_number + (1|Hopper), data = escape)
summary(model)


#estimate repeatability by dividing the amount of variation explained by individual ID by the total variance explained by individual ID and residuals (VID / VID + VR)

var_components <- as.data.frame(VarCorr(model))
var_components

VID <- var_components$vcov[var_components$grp == "Hopper"]
VID

VR <- attr(VarCorr(model), "sc")^2
VR

repeatability <- VID / (VID + VR)
repeatability
```


```{r}
table_contact_location <- table(escape$Contact, escape$Location)
table_contact_location

#Chi-Square test
chisq_test <- chisq.test(table_contact_location)
print(chisq_test)

#Fisher's Exact test
fisher_test <- fisher.test(table_contact_location)
print(fisher_test)

```
```{r}
a <- ggplot(escape, aes(x = Location, y = Initiation_distance, fill = Location)) +
geom_violin (trim = FALSE, alpha = 0.5) +
  geom_jitter(color = "black", width = 0.15, alpha = 0.5) +
  scale_fill_manual(values=c("grey", "grey")) +
  labs(x = "", y = "Initiation Distance (cm)") +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.title.x = element_text(margin = margin(t = 18), size = 0),  
        axis.title.y = element_text(margin = margin(r = 10), size = 12),
        axis.text.x = element_text(margin = margin(t = 4), size = 12, color = "black"))

a

ggsave("initiation_distance.pdf", path = "figures/")
ggsave("initiation_distance.png", path = "figures/")

b <- ggplot(escape, aes(x = Location, fill = factor(Contact))) +
  geom_bar_pattern(
    aes(pattern = factor(Contact)), width = 0.8,  
    position = "fill",  
    pattern_density = 0.1,  
    pattern_angle = 45,  
    pattern_spacing = 0.05,  
    color = "black",  
    size = 0.5 
  ) +
  scale_pattern_manual(values = c("0" = "circle", "1" = "crosshatch")) +  
  scale_fill_manual(values = c("0" = "white", "1" = "white")) +
  labs(y = "Deimatic display initiated\nby physical contact", x = "Location", fill = "Contact") +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.title.x = element_text(margin = margin(t = 18), size = 0),  
        axis.title.y = element_text(margin = margin(r = 10), size = 12),
        axis.text.x = element_text(margin = margin(t = 4), size = 12, color = "black"))

b

ggsave("contact_proportion.pdf", path = "figures/")
ggsave("contact_proportion.png", path = "figures/")

grid.arrange(a, b, ncol = 2)

a <- a + plot_annotation(tag_levels = "A")
B <- b + plot_annotation(tag_levels = "B")

figure_2 <- a + b

figure_2

ggsave("Figure_2.pdf", path = "figures/")
ggsave("Figure_2.png", path = "figures/")

#touch 20
#non-touch 11
#total 31

20/31
```

#the first flight for each grasshopper represents their initial location and doesn't reflect the true choice of background (since it's the starting point). I need to remove the first flight for each grasshopper before performing a glmm
```{r}
# Remove the first flight for each hopper
clean_data <- escape %>%
  group_by(Hopper) %>%
  filter(Flight_number > 1) %>%
  ungroup()

glmm_model <- glmer(Location ~ 1 + (1 | Hopper), 
                    data = clean_data, 
                    family = binomial)

# View the model results
summary(glmm_model)
```


To test if the distance traveled decreases after each jump, you can use a linear mixed-effects model (LMM) because your data has multiple measurements per grasshopper (i.e., repeated measures).
```{r}
data_clean <- escape[!is.na(escape$Distance_traveled), ]

data_clean

# Fit a linear mixed effects model
model <- lmer(Distance_traveled ~ Flight_number + (1 | Hopper), data = data_clean)

# Summarize the model
summary(model)
```


```{r}
R.version
citation()
```

